---
title: "Ishihama H-NS gSELEX data"
author: "Mike Wolfe"
date: "3/20/2022"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE, fig.align = "center")
```


```{r}
library(tidyverse)
theme_set(theme_classic())
```

We wanted to look at whether some of the sequence features that were associated
with our footprinting data were also associated with the Ishihama H-NS gSELEX
data.

To perform this analysis, I took the peak centers from Table S1 from their paper
(https://doi.org/10.1093/femsre/fuab032).

It was unclear from this table or their paper, which *E. coli* K-12 genome their
coordinates referred to. To determine which coordinates made sense, I checked
their coordinates and their corresponding annotations for each peak against the
U00096.3 and U00096.2 genomes. By matching their peak annotations against each
genome, it appears they used a U00096.2 genome rather than the more recent
U00096.3 genome.

Since, they do not denote which genome their coordinates are in, this is a best
guess but manually checking a handful of sites at the end of the genome (where
coordinates are most impacted by the extra large insertions in U00096.3 compared
to U00096.2), I can be pretty confident that these are the right sites.

I next took a window of 50 bp on either side of the peak center for a total of
101 bp for each peak. To make a null set, I randomly sampled the genome for
a window of the same size for each H-NS peak.

I then used the same scripts I have been using for everything else to calculate
the locations of TA-steps, A-tracts, the value of a 5 bp windowed A/T content,
and DNA shape parameters.

## Sequence features

First, I wanted to see if TA-steps or A-tracts were enriched in these H-NS peaks
compared to the null set.

```{r, include = FALSE}
d <- read_tsv("tidy_seq_data.tsv.gz")
head(d)
```

```{r, include = FALSE}
d <- d %>% mutate(Site = if_else(str_detect(contig, "RAND", negate = TRUE), "H-NS bound", "Random sites")) %>%
    mutate(value = value > 0)
head(d)
```

```{r, fig.height = 5, fig.width = 7}
d %>% 
    group_by(param, contig, Site) %>%
    summarize(frac_count = sum(value > 0)/n()) %>%
    ggplot(aes(x = frac_count, fill = Site)) + geom_histogram(position = "identity", alpha = 0.5) +
    geom_vline(data = . %>% group_by(param, Site) %>% summarize(mean = mean(frac_count, na.rm = TRUE)),
               aes(xintercept = mean, color = Site), linetype = "dashed") +
    facet_wrap(~param, scales = "free_y", ncol = 1) +
    scale_fill_manual(values = c("#DA291CFF", "grey40")) +
    scale_color_manual(values = c("#DA291CFF", "grey40")) +
    labs(x = "Fraction of bp in each feature", title = "Enrichment of Sequence Features in H-NS gSELEX sites",
         subtitle = "Dashed lines indicated means")
```

We can test for this enrichment using a permutation test by using the difference
in averages between the two classes as a test statistic, shuffling the labels
10,000 times, and asking how many shuffled sets had a larger difference in means
than the actual data. This is a two-sided permutation test.
```{r}
perm_test <- function(values, labels, n = 10000) {
    actual <- abs(mean(values[labels]) - mean(values[!labels]))
    tests <- numeric(n)
    for(i in 1:n){
        shuff_labels <- sample(labels, size = length(labels), replace = FALSE)
        tests[i] <- abs(mean(values[shuff_labels]) - mean(values[!shuff_labels]) >= actual)
    }
    return((sum(tests) + 1)/ (n +1))
}
        
```

```{r}
d %>% 
    group_by(param, contig, Site) %>%
    summarize(frac_count = sum(value > 0)/n()) %>%
    group_by(param) %>%
    summarize(perm_pval = perm_test(frac_count, Site == "H-NS bound"))
```

The permutation test indicates that both features appear to be enriched in H-NS
sites relative to random locations.


## General A/T content

We can also look at the A/T content of the sites. Here I took a 5 bp window to
look at the A/T content across each site just like I did in the paper.
```{r, include = FALSE}
d_at <- read_tsv("Ishihama_windowed_at.tsv.gz")
head(d_at)
```

```{r, include = FALSE}
d_at <- d_at %>% mutate(Site = if_else(str_detect(contig, "RAND", negate = TRUE), "H-NS bound", "Random sites")) 
head(d_at)
```


First we can look at the overall distribution of each sites 5 bp A/T window
content

```{r}
d_at %>% group_by(Site, contig) %>%
    summarize(avg_at = mean(value)) %>%
    ggplot(aes(x = avg_at, fill = Site)) + geom_histogram(position = "identity", alpha = 0.5) +
    geom_vline(data = . %>% group_by(Site) %>% summarize(mean = mean(avg_at, na.rm = TRUE)),
               aes(xintercept = mean, color = Site), linetype = "dashed") +
    geom_text(data = . %>% group_by(Site) %>% summarize(mean = mean(avg_at, na.rm = TRUE)) %>%
                  mutate(adjusted = if_else(Site == "H-NS bound", mean + 0.05, mean - 0.05)),
              aes(x = adjusted, y = 125, label = sprintf("%.2f", mean), color = Site)) +
    scale_fill_manual(values = c("#DA291CFF", "grey20")) +
    scale_color_manual(values = c("#DA291CFF", "grey20")) +
    labs(x = "Average of A/T content over 5bp windows", title = "Enrichment of Sequence Features in H-NS gSELEX sites",
         subtitle = "Dashed lines indicated means")
```

We can get a p-value for the difference in these means using a permutation test
with shuffled labels like I had above.
```{r}
d_at %>% 
    group_by(contig, Site) %>%
    summarize(avg_at = mean(value)) %>%
    ungroup() %>%
    summarize(perm_pval = perm_test(avg_at, Site == "H-NS bound"))
```
The H-NS sites are clearly enriched for 5 bp windowed A/T content over the
random sites which is not too surprising.


## Shape parameters

I also wanted to look at the SHAPE data in a similar manner

```{r, include = FALSE}
ds <- read_tsv("Ishihama_shape_tidy.tsv.gz")
head(ds)
```


```{r, include = FALSE}
ds <- ds %>% mutate(Site = if_else(str_detect(contig, "RAND", negate = TRUE), "H-NS bound", "Random sites"))
head(ds)
```
First let's look at any sort of overall enrichment in either direction for each
of these SHAPE parameters

```{r}
ds %>% group_by(param, contig, Site) %>%
    summarize(avg_value = mean(value, na.rm = TRUE)) %>%
    ggplot(aes(x = avg_value, fill = Site)) + geom_histogram(position = "identity", alpha = 0.5) +
    geom_vline(data = . %>% group_by(Site, param) %>% summarize(mean = mean(avg_value, na.rm = TRUE)),
               aes(xintercept = mean, color = Site), linetype = "dashed") +
    facet_wrap(~param, scales = "free") +
    scale_fill_manual(values = c("#53A567FF", "grey40")) +
    scale_color_manual(values = c("#53A567FF", "grey40")) +
    labs(title = "Enrichment of each SHAPE feature",
         x = "Average value of parameter in each site",
         subtitle = "Dashed lines indicate means")
        
```

Now lets do a two-sided permutation test for these features. 

```{r}
ds %>% 
    group_by(contig, Site, param) %>%
    summarize(avg_val = mean(value, na.rm = TRUE)) %>%
    group_by(param) %>%
    summarize(perm_pval = perm_test(avg_val, Site == "H-NS bound"))
```
Each of thes shape features show a statistically significant difference between
the H-NS sites and the random sites. However, the magnitude of these changes is
pretty small for some of the SHAPE features. For example, Roll clearly seems to
just have higher variation in values in the H-NS sites compared to the random
sites rather than a shift in the average from looking at the distributions.


## Putting it all together for the paper.

The general conclusions I get out of this analysis is that each of the
features we identified in our logistic regression model as important
for H-NS DBD binding seem to be enriched in the gSELEX sites compared to
random sites across the genome.

Here, I put together the most relevant information from this analysis into a
figure we use for the paper. I chose the features that were chosen
by our LASSO regression to focus on.

```{r}
library(patchwork)
```



```{r}
ta_den <- d %>% filter(param == "TA_step") %>%
    group_by(Site, contig) %>% summarize(frac_count = sum(value > 0) / n()) %>%
    ggplot(aes(x = frac_count, fill = Site)) + geom_histogram(alpha = 0.5, position = "identity") +
    geom_vline(data = . %>% group_by(Site) %>% summarize(mean = mean(frac_count, na.rm = TRUE)),
               aes(xintercept = mean, color = Site), linetype = "dashed") +
     theme_classic() +
     theme(legend.position = "none") +
     scale_fill_manual(values = c("#DA291CFF", "grey40")) +
     scale_color_manual(values = c("#DA291CFF", "grey40")) +
    labs(x = "Fraction of bp in TA step per site", title = "TA step")

at_den <- d_at %>% filter(param == "A/T_5bp") %>%
    group_by(Site, contig) %>% summarize(avg_at = mean(value)) %>%
    ggplot(aes(x = avg_at, fill = Site)) + geom_histogram(alpha = 0.5, position = "identity") +
    geom_vline(data = . %>% group_by(Site) %>% summarize(mean = mean(avg_at, na.rm = TRUE)),
               aes(xintercept = mean, color = Site), linetype = "dashed") +
     theme_classic() +
    theme(legend.position = "none") +
     scale_fill_manual(values = c("#DA291CFF", "grey40")) +
     scale_color_manual(values = c("#DA291CFF", "grey40")) +
    labs(x = "Average 5 bp A/T content per site", title = "A/T content")

mgw_den <- ds %>% filter(param == "MGW") %>%
    group_by(Site, contig) %>% summarize(avg = mean(value, na.rm = TRUE)) %>%
    ggplot(aes(x = avg, fill = Site)) + geom_histogram(alpha = 0.5, position = "identity") +
    geom_vline(data = . %>% group_by(Site) %>% summarize(mean = mean(avg, na.rm = TRUE)),
               aes(xintercept = mean, color = Site), linetype = "dashed") +
     theme_classic() +
    theme(legend.position = "none") +
     scale_fill_manual(values = c("#53A567FF", "grey40")) +
     scale_color_manual(values = c("#53A567FF", "grey40")) +
    labs(x = "Average MGW (Å) per site", title = "Minor Groove Width (Å)")

ep_den <- ds %>% filter(param == "EP") %>%
    group_by(Site, contig) %>% summarize(avg = mean(value, na.rm = TRUE)) %>%
    ggplot(aes(x = avg, fill = Site)) + geom_histogram(alpha = 0.5, position = "identity") +
    geom_vline(data = . %>% group_by(Site) %>% summarize(mean = mean(avg, na.rm = TRUE)),
               aes(xintercept = mean, color = Site), linetype = "dashed") +
     theme_classic() +
    theme(legend.position = "none") +
     scale_fill_manual(values = c("#56A8CBFF", "grey40")) +
     scale_color_manual(values = c("#56A8CBFF", "grey40")) +
    labs(x = "Average EP (kT/e) per site", title = "Electrostatic Potential (kT/e)")
```

For these plots, the gray is the random sites and the colors are the H-NS sites.

```{r, fig.width = 8, fig.height = 5}
(ta_den + at_den + mgw_den + ep_den) + plot_annotation(tag_levels = 'A', tag_suffix= '.')
```

```{r}
ggsave("Ishihama_feature_enrichment.pdf", height = 5.0, width = 8.0, units = "in")
```

    
Get averages the requisite averages as well
```{r}
d_at %>% filter(param == "A/T_5bp") %>%
    group_by(Site, contig) %>% summarize(avg_at = mean(value)) %>%
    group_by(Site) %>%
    summarize(mean(avg_at))
```


```{r}
ds %>% filter(param == "MGW") %>%
    group_by(Site, contig) %>% summarize(avg = mean(value, na.rm = TRUE)) %>%
    group_by(Site) %>%
    summarize(mean(avg))
```

```{r}
ds %>% filter(param == "EP") %>%
    group_by(Site, contig) %>% summarize(avg = mean(value, na.rm = TRUE)) %>%
    group_by(Site) %>%
    summarize(mean(avg))
```

```{r}
d %>% filter(param == "TA_step") %>%
    group_by(Site, contig) %>% summarize(frac_count = sum(value > 0) / n()) %>%
    group_by(Site) %>%
    summarize(mean(frac_count))
```


