```{r}
library(tidyverse)
theme_set(theme_classic())
```


# Turning the match scores into p-values

Need an offset function to make the center of the DBD in the MD pattern
```{r}
offset <- function(data){
  data %>% mutate(bp = ifelse(direction == "fwd", bp-2, bp + 2))
}
```


Read in the match scores for each modality and convert first to original
coordinate system
```{r}
FEbabe <- read_tsv("../md_sim_pattern_matching/model_match_1shuff_match_scores.tsv") %>%
    mutate(method = "Tethered") %>%
    mutate(bp = bp + 643) %>%
    offset()
OH <- read_tsv("../md_sim_pattern_matching/OH_model_match_1shuff_match_scores.tsv") %>%
    mutate(method = "OH") %>%
    mutate(bp = bp + 643) %>%
    offset()
d <- bind_rows(FEbabe, OH)
```
Now let's pull in the shuffled data
```{r}
(FEbabe_shuff <- read_tsv("../md_sim_pattern_matching/model_match_1shuff_match_scores_rand.tsv") %>%
    mutate(bp = bp + 643) %>%
     separate(direction, sep = "_", into= c("direction", "sample")) %>%
     offset()
)
```

```{r}
(OH_shuff <- read_tsv("../md_sim_pattern_matching/OH_model_match_1shuff_match_scores_rand.tsv") %>%
     mutate(bp = bp + 643) %>%
     separate(direction, sep = "_", into= c("direction", "sample")) %>%
     offset()
)
```

     
Now we define a function to get the permutation-based pvalue. We did 1000 
samples and we will add a psuedocount of one to the top and the bottom of the
denominator to deal with the scores more extreme than all shuffles.

(This is like a prior where you expect 50-50)
```{r}
get_p_value <- function(real, sampled){
    pvals <- list()
    for(i in seq_along(real$bp)){
        true_val <- real$match_score[[i]]
        this_pval <- sampled %>% filter(real$bp[[i]] == bp & real$direction[[i]] == direction) %>%
            summarize(total = sum(match_score <= true_val)) %>%
            pull(total)
        pvals[[i]] <- (this_pval + 1)/(1000 +1)
    }
    real %>% mutate(pval = unlist(pvals))
}
```

Get the pvalues for each and also convert the coordinate system to the
new one
```{r}
(FEbabe <- get_p_value(FEbabe, FEbabe_shuff) %>%
    mutate(bp = abs(bp - 756)) %>%
    mutate(direction = if_else(direction == "fwd", "Reverse", "Forward")))
```

```{r}
OH <- get_p_value(OH, OH_shuff) %>%
    mutate(bp = abs(bp - 756)) %>%
    mutate(direction = if_else(direction == "fwd", "Reverse", "Forward"))
```


Now plot each individual p-value trace and the combined p-value trace
```{r}
(p1 <- FEbabe %>%
    ggplot(aes(x = bp, y = -log10(pval), color = direction)) + geom_point() + geom_line() +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
    scale_color_manual(values = c("#009445", "#1C75BC")) +
    labs(x = "DRE position (bp)", y = "-log10(pvalue)", title = "TEN-map & MD agreement",
         color = "DBD\nOrientation"))
```

```{r}
tethered_cutoffs <- tibble(md_percentiles = 
        factor(c(">50%", "<50%", "<10%","<5%", "<2.5%","<1%"), 
               levels = c(">50%", "<50%", "<10%","<5%", "<2.5%","<1%")),
        cutoffs = c(Inf, 
                    quantile(FEbabe_shuff$match_score, 0.50),
                    quantile(FEbabe_shuff$match_score, 0.10),
                    quantile(FEbabe_shuff$match_score, 0.05),
                    quantile(FEbabe_shuff$match_score, 0.025),
                    quantile(FEbabe_shuff$match_score, 0.01)))
tethered_cutoffs <- tethered_cutoffs %>% mutate(span = lead(cutoffs))
head(tethered_cutoffs)
```


```{r}
FEbabe %>% 
   ggplot(aes(x = bp, y = match_score, color = direction)) +
   geom_rect(data = tethered_cutoffs,
             inherit.aes = FALSE,
             aes(ymin = span, ymax = cutoffs, 
                 fill = md_percentiles),
             xmin = -Inf, xmax = Inf, alpha = 0.8, color = NA) +
   scale_fill_brewer(palette = "Greys", direction = -1,
                     guide = guide_legend(ncol=2,
                                          title.position="top",
                                          label.position="right")) +
   geom_point(size = 1) + 
  geom_line(size = 1) + theme_bw() + 
     scale_color_manual(values = c("#009445", "#1C75BC"),
                        guide = guide_legend(ncol=1, 
                                             title.position = "top",
                                             label.position = "right")) + 
     geom_hline(yintercept = 
                  tethered_cutoffs$cutoffs[tethered_cutoffs$md_percentiles == "<5%"],
                linetype="dashed") + 
     labs(title = "TEN-map", y = "MD Match Score",
          color = "DBD Orientation", x= "DRE position (bp)",
          fill = "Shuffled Match Score\nPercentiles") +
   theme(aspect.ratio = 0.3, legend.position = "right")
```

```{r}
ggsave("tenmap_md_percentiles_plot.pdf")
```

```{r}
(p2 <- OH  %>%
    ggplot(aes(x = bp, y = -log10(pval), color = direction)) + geom_point() + geom_line() +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
    scale_color_manual(values = c("#009445", "#1C75BC")) +
    labs(x = "DRE position (bp)", y = "-log10(pvalue)", title = "Footprinting & MD agreement",
         color = "DBD\nOrientation"))
```



Combine the two pvalues using Fisher's method
```{r}
d <- rbind(FEbabe, OH)
```
Get the locations where either method is less than 0.05
```{r}
d %>% group_by(bp, direction) %>%
  summarize(binding = sum(pval < 0.05)) %>%
  mutate(bp = 145 - bp) %>%
  write_tsv("combined_individ_binding.tsv")
```

```{r}
(p3 <- d %>% group_by(bp, direction) %>%
    # fishers method for combining pvalues
    summarize(combined_p = 1-pchisq(-2*sum(log(pval)),4)) %>%
    ggplot(aes(x = bp, y = -log10(combined_p), color = direction)) + geom_point() +
    geom_line() +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed") + 
    scale_color_manual(values = c("#009445", "#1C75BC")) +
    labs(x = "DRE position (bp)", y = "-log10(pvalue)", title = "Combined p-values using Fisher's method",
         color = "DBD\nOrientation"))
```


```{r, fig.width = 7, fig.height = 7}
library(patchwork)
p1/p2/p3
```

```{r}
ggsave("match_score_p_value_plots.pdf")
```
